---
title: "Data Science Log #3"
author: "Rrustem Sallauka"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Pixar Films

For small datasets like these adding them to git is not a problem.
```{r}
library(tidyverse)
```

    Load the `pixar_films` dataset into R. Clean the dataset by removing
    films without a title. Inspect the variable `film_rating`. What are
    the possible values and what do they mean? Create a factor variable
    for the film rating. Why is this appropriate?
```{r}
pixar_films <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/pixar_films.csv')
pixar_films <- pixar_films %>% drop_na(film)
pixar_films$film_rating <- factor(pixar_films$film_rating, exclude = NULL)
```

The possible values are G and PG. The rating based on Motion Picture Association (MPA) film rating system. 
G: General Audiences
PG: Parental Guidance Suggested

film_rating is a categorical variable,because it has a fixed and known set of possible values, which when we transform can help us analising and vizualising
<!-- -->

    Inspect the film titles manually. Which films form a film series? A
    film series can be identified by a common word in the titles of the
    films, often in conjunction with a number in the title,
    e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which
    displays a list of the different series with the titles of the films
    and how many films belong to the series. Output the dataframe.

Cars, Toy Story, Incredibles, Finding Nemo 
```{r}
film_series <- pixar_films %>%
  filter(str_detect(film, "Cars") | 
         str_detect(film, "Toy Story") | 
         str_detect(film, "Incredibles") | 
         str_detect(film, "Finding")| 
         str_detect(film, "Monsters")) %>% mutate(
           series = case_when(
         str_detect(film, "Cars") ~ "Cars" , 
         str_detect(film, "Toy Story") ~ "Toy Story", 
         str_detect(film, "Incredibles") ~ "Incredibles", 
         str_detect(film, "Finding") ~ "Finding", 
         str_detect(film, "Monsters") ~ "Monsters")
         )
film_series <- film_series %>% group_by(series) %>% mutate(count = n()) 

film_series
```


    Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.
```{r}
public_response <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-11/public_response.csv')

public_response$cinema_score <- factor(public_response$cinema_score, levels = c("A-","A","A+"), exclude = NULL, ordered = TRUE)
public_pixar <- merge(film_series, public_response, by = "film")
```


    Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?
```{r}
ggplot(public_pixar <- public_pixar %>% mutate(release_date = as.Date(release_date)),
  aes(x = reorder(film, release_date), y = critics_choice, fill = film)) +
  geom_col(position = "dodge") +
  facet_wrap(~series, scales = "free_x") +
  labs(
    title = "Critics Choice for Pixar Film Series",
    x = "Film Title",
    y = "Critics Choice Score"
  ) +
  theme_minimal()
```

The trends are similar, the first movie is ranked better by the critics then the following movies, in some series like Cars the score goes down. 

## Open Analysis 

```{r}
energy_prices <- read_csv("data/Energy Prices - BP Statistics (2016).csv")
```

Put the name / title of the dataset and a link to it below.

-   Dataset Name: Energy Prices - BP Statistics (2016)
-   Link: <https://github.com/owid/owid-datasets/tree/master/datasets/Energy%20Prices%20-%20BP%20Statistics%20(2016)#energy-prices---bp-statistics-2016>

    Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
    How have energy prices changed over time?
    
    Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.

```{r}
energy_prices <- energy_prices %>% pivot_longer(cols = -c(Entity,Year),names_to = "source", values_to = "price")

energy_prices
```    
    
    Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.
```{r}
ggplot(
  energy_prices,
  aes(
    x = Year,
    y = price,
    colour = source
  )
) + geom_line() + labs(
  title = "How have energy prices changed over time?",
  x = "Year",
  y = "Price",
  color = "Source"
)
```
